\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-09 paper/experiments.tex

\section{Experiments}
\todo{we of course need to rewrite this whole section}
To evaluate the performance of our system, we generate plans for queries on two data sets: the American Community Survey and a synthetic dataset.
\subsection{Data sets} \label{subsec:datasets}

The American Community Survey (ACS) provides a number of public data sets collected by the U.S. Census Bureau.
We used a cleaned version of the 2012 Public Use Microdata Sample (PUMS) data kindly provided by the authors of~\cite{akande2015empirical}.
The cleaning procedure is described in detail in~\cite{akande2015empirical}, but to summarize, the following data were removed:
\begin{itemize}
\item Rows corresponding to vacant houses or single occupant households.
\item Identification variables (e.g. area codes, serial numbers).
\item Flag variables.
\item Continuous variables.
\end{itemize}
The final data set consists of a single table with 671,153 rows and 37 columns, where all column entries are integers.

In order to evaluate our implementation, we created a dirtied variant of the ACS data by randomly eliminating
10\% of the fields across the table.

In order to evaluate additional queries, we constructed a synthetic table, consisting of 10,000 rows and 10 columns, with
all values drawn from a uniform distribution between 0 and 100. To create a dirtied variant of the synthetic data, we randomly deleted 30\% of the fields
across the table.

\subsection{Queries}

We collected a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.

The queries on the ACS data all consist only of projections, selections, and aggregations.
The ACS data is contained in a single table, so there are relatively few join queries that are interesting on this data set and remain relevant for analysts, rather
than contrived solely for the purpose of experimentation.
However, even with select, filter, and aggregate, there are interesting choices to make with imputation placement. In order to explore joins, we leverage the synthetic
dataset and write various ad-hoc queries.

>>>>

\section{Experiments}
\todo{we of course need to rewrite this whole section}
To evaluate the performance of our system, we generate plans for queries on two data sets: the American Community Survey and a synthetic dataset.
\subsection{Data sets} \label{subsec:datasets}
\subsubsection{Center for Disease Control: Surveys}
For our first set of experiments, we use survey data collected by the 
Center for Disease Control (CDC) in the United States. In particular, we
experiment on a set of tables collected as part of the National
Health and Nutrition Examination Survey (NHANES), a series of studies
conducted by the CDC on a national sample of several thousand individuals\cite{nhanes}.
The data consists on survey responses, physical examinations, and laboratory
results, amongst others. Our dataset corresponds to the 2013-2014 NHANES
and is available for direct download on the CDC's organization site on 
kaggle.com, a popular data science website.

There are a total of 6 tables associated with the NHANES data set. For purposes
of our experiments, we use three tables:

\begin{itemize}
	\item Demographics: outlines demographic information of subjects
	\item Examinations: physical exam results
	\item Laboratory: laboratory exam results
\end{itemize}

The original tables have a large number of attributes, in some cases providing more granular
tests results or alternative metrics. In order to simplify the presentation and
exploration of queries, we focused on a subset of the attributes for each table.
Table~\ref{table:nhanes-description} shows the attributes selected, along with the
percentage of nulls. It is worth noting that for readability, we have replaced the
NHANES variable names with self-explanatory attribute names.

\begin{table}
    \centering
    \begin{subtable}{0.5\textwidth}
        \centering
       \input{tables/cdc_demo}
        \caption{Demographics}
    \end{subtable}
    \begin{subtable}{0.5\textwidth}
        \centering
        \input{tables/cdc_labs}
        \caption{Laboratory Results}
    \end{subtable}
       \begin{subtable}{0.5\textwidth}
        \centering
        \input{tables/cdc_exams}
        \caption{Physical Results}
    \end{subtable}
    \caption{Missing values in CDC NHANES 2013-2014 data}
\end{table}

All tables used consists of 10 attributes. The laboratory and examinations data,
referred to as \textit{labs} and \textit{exams} in our queries, consists of 9813 rows,
while the demographics data (\textit{demo}) has 10175 rows.

\subsubsection{Some Other TBD dataset}
\todo{write data description for this}

\subsection{Queries}
We collected a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.

The queries on the CDC NHANES data consists not only of projections and selections,
but also interesting joins and aggregations. Our aim was to craft meaningful queries
that would provide performance figures relevant to practitioners who may use
similar datasets.

Query~\ref{q1} calculates the average weight for users based on their income data.
To do so, it joins demographics and physical examination data. Query~\ref{q2}
compares cholesterol levels for individuals who are both ends of the income
spectrum and who satisfy a minimum weight requirement. Query~\ref{q3}
extracts the maximum blood lead levels for subjects who satisfy a particular
waist circumference requirement.

\todo{describe further queries on other data set}
